Preliminary Draft

### 2021-07-19 article/draft.tex
minor changes

@@

A number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.''}

Section $230$ in the United States Communications Decency Act\footnote{Similar regulation exists in the European Union, see articles $12$ and $15$ of the E-commerce Directive (2000). } provides immunity for website platforms against the content created by users. Nevertheless, there is growing pressure for Mainstream Social Media Platforms (hereafter MSMP), such as Facebook, Twitter or Youtube, to moderate the available content. In particular, platforms seem to take explicit actions when content is in violation of local laws in different jurisdictions, e.g. laws regarding defamation of a racial nature, dissemination of symbols from unconstitutional organizations, privacy protection, digital security, electoral laws. Facebook reports having implemented a total of $64.7$ thousand content restrictions based on local law across all countries in 2020.\footnote{See Facebook Transparency Center, Content restrictions based on Local Law: \href{https://transparency.fb.com/data/content\-restrictions}{transparency.fb.com/data/content\-restrictions}. We summed the count of content restrictions over all countries reported in the table, for $H1$ and $H2$ of the year 2020.} Google reports a total of $26$ thousand government requests to remove content from July 2020 to December 2020, among which $11.4$ thousand concerned Youtube.\footnote{See Google's Transparency report, government requests to remove content: \href{https://transparencyreport.google.com/government-removals/overview}{transparencyreport.google.com/government-removals/overview}.} Twitter reports having received $42.2$ thousand legal demands from third-parties from January to June 2020, and has responded by withholding $82$ thousand accounts and $3.1$ thousand tweets.\footnote{See Twitter Transparency website, Removal requests: \href{https://transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}{transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}.}

@@

Furthermore, MSMP are increasingly engaging in editorial tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and Youtube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam.\footnote{For an exhaustive overview of the community standards of Facebook, see: \href{https://www.facebook.com/communitystandards/}{facebook.com/communitystandards/}. For the Twitter  Rules see: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{help.twitter.com/en/rules-and-policies/twitter-rules}, and for Youtube community guidelines see: \href{https://www.youtube.com/intl/en\_us/howyoutubeworks/policies/community\-guidelines/}{youtube.com/intl/en\_us/howyoutubeworks/policies/community\-guidelines/}.} While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries.

@@

In this article, we focus on MSMP's policies and actions regarding content with low credibility or false information, commonly referred to as {\it Fake News}.\footnote{For an overview on the {\it concept} of {\it Fake News}, we refer the reader to the article The science of fake news by Lazer et al. (2018)~\cite{lazer}.} The {\it Fake News} phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience. The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society, in the rare event that it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. \footnote{For an overview of Facebook's third-party fact checking program see: \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}.} Facebook use a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section} Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation.\footnote{See Twitter Birdwatch: \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}.}

@@

{\color{pink} Note, put here something along the following lines: cite misinformation review article about support for fact-checking + hard to correct beliefs once it goes viral. cite Pennycook / Rand? + where to put or cite \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{this} ? }
%In particular, determining the truthfulness of content over controversial topics or ongoing political events proves to be a tricky task and correcting beliefs about viral content, fact-checked as false ... 
%For an overview of Youtube's policy on fact-checking see: \href{https://support.google.com/youtube/answer/9229632}{support.google.com/youtube/answer/9229632}. To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.}

%In particular, the peculiar task of determining the truthfulness of content over controversial topics or ongoing political events 
\smallskip

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content.\footnote{For Facebook, Twitter and Youtube see respectively the following updates concerning COVID19 related misinformation policies: \href{https://support.google.com/youtube/answer/9891785}{https://support.google.com/youtube/answer/9891785}, \href{https://help.twitter.com/en/rules-and-policies/medical-misinformation-policy}{help.twitter.com/en/rules-and-policies/medical-misinformation-policy} and \href{https://support.google.com/youtube/answer/9891785}{support.google.com/youtube/answer/9891785}.} Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged\footnote{{\color{pink} put here the list of Facebook and Twitter terms about covid.}}, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of MSMP to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines\footnote{As of July 2021.}. The 2019 report of the Facebook Data Transparency Advisory Group\footnote{To read the full report, see: \href{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report?fbclid=IwAR2xMZr5GdD1GaNpjsXR3\_yeeIR4H9iFASfrni5HKcJVAO5oWA52bvwcZxU}{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report?fbclid=IwAR2xMZr5GdD1GaNpjsXR3\_yeeIR4H9iFASfrni5HKcJVAO5oWA52bvwcZxU}.} (DTAG) states clearly that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report.}''. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom.\footnote{See: \href{https://about.fb.com/news/2018/05/hard-questions-false-news/}{https://about.fb.com/news/2018/05/hard-questions-false-news/}.}. Similarly, Twitter communicates about actions related to misinformation via their Twitter Safety Blog/account\footnote{See \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{https://blog.twitter.com/en\_us/authors.TwitterSafety}.}. In a post on Youtube Official Blog\footnote{See \href{https://blog.youtube/inside-youtube/the-four-rs-of-responsibility-raise-and-reduce/}{https://blog.youtube/inside-youtube/the-four-rs-of-responsibility-raise-and-reduce/}.}, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content reduces borderline content and harmful misinformation.

{\color{pink} Note, put here something along the following line: make point here about the fact that in the transparency centers we find things for the violations in the platforms rules but not misinformation.... needed so that the academic community can study the reach and impact of this phenomena and provide guidelines for adequate actions... + where to put this \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} ? }
%{\color{pink}  idea about academic community needs to be able to study this phenomenon and assess the impact of the policies, what effects they have etc. Violations fall into other categories: deceptive practices, impersonation, hate speech.  }

@@

In the present article, we will explain how to verify with data mining MSMP's actions regarding content with low credibility or false information, through a series of examples for different actions and platforms. For the purpose of clarity, we only focus on three platforms: Facebook, Twitter and Youtube. Both Facebook and Youtube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. More specifically, we survey a number of common policies used in order to tackle misinformation, across the three above cited MSMP, Facebook, Twitter and Youtube: temporary or permanent suspension of users, reducing the visibility of some content, introducing flags and notices. We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies. We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. We summarize in table \ref{tab1} the tools used to collect the data from Facebook, Twitter and Youtube, that we use in multiple examples throughout the present article. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.  %the announced policy for a specific user was implemented or not.

@@

When an account is permanently suspended by Facebook, it disappears from the platform. That is,  the data can no longer be scrapped and it also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.} Facebook publishes on monthly basis a {\it coordinated inauthentic behavior} report, where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged.\footnote{See the \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{April 2021 report} as an example.} But as long as external persons do not have access to deleted accounts data, these reports cannot be verified by independent researchers or journalists.

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. Thus, after manually adding this page to the CrowdTangle dashboard, we collected the $6$ $083$ posts it had published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{See the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used the minet command line tool \cite{minet} to collect data.\footnote{The exact command can be found \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_trump_data.sh}{here}.} We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

@@

The website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$}; which was recently suspended.\footnote{We noticed the message about the account suspension on May 25, 2021. But to the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended anytime between April 15, 2021 and May 25, 2021 (see the suspension screenshot in panel $(a)$ of figure \ref{fig3}). Furthermore, $globalresearch.ca$ has multiple failed fact-checks, see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/}} When a user searches via the twitter search-box for any URL link of this website, no results appear as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, taken on June $14$, $2021$. To further investigate the possible implementation of a reduced visibility measure, we search via the Twitter API for tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$.  As shown in panel (a) in figure \ref{fig4}, we find a strictly positive number of tweets containing the URL link {\it globalresearch.ca} throughout May $2021$ and the first week of June $2021$. Hence, the visibility of tweets containing this URL link has been reduced because users can no longer access tweets containing the URL link {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this URL, as shown in the screenshot in panel $(a)$ of figure \ref{fig4bis}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who tweet articles from the {\it globalresearch.ca} website receive tweet level engagement from their own followers. Finally, when a user attempts to click on the URL link {\it globalresearch.ca} contained in the Tweet, a warning message appears and indicates that the link may be unsafe (see screenshot in panel (b) in figure \ref{fig4bis}).

@@

We further examine the placement of interstitials that indicate a possibly sensitive content. We find that only $2.97\%$ out of $323$ $938$ tweets containing a URL marked as False, have an interstitial ``potentially sensitive content''. In particular, many speak about COVID19 and do not contain a label to provide users with more context from authoritative sources. Figure \ref{fig_notice} provides an example of a tweet sharing a URL marked as False by a Fact-checker and which contains an interstitial ``potentially sensitive content''. We find 32 other Tweets, among our set of collected tweets, who share the exact same URL link as in the previous example. Among those 32 Tweets, only 5 tweets had the interstitial ``potentially sensitive content''. Again this points towards the non-automation of the interstitials placement.

@@

Finally, the Twitter Safety announced in a Tweet\footnote{See Tweet ID : \href{https://twitter.com/TwitterSafety/status/1379515615954620418}{1379515615954620418}} on April 6, 2021 that their team ``will begin deploying automated tools to build on (their) efforts to label tweets that may contain misleading information around COVID-19 vaccinations". 
%we don't know what it means the articl xxx which was retracted by the journal and fack-check still appears in xx tweets with no label

@@

@@

%https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works
\end{document}

>>>>

Title proposition 3: Misinformation Policies of Mainstream Social Media Platforms: third-party verification methods and illustration

@@

A number of recent studies  point towards the idea that ``Fake News'' or disinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense. \footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.'' Furthermore, the European commission recognizes the spread of online disinformation as a problem and has put together in 2018 a \href{https://digital-strategy.ec.europa.eu/en/policies/code-practice-disinformation}{Code of practice on Disinformation}, which is a set of self-regulatory standards to fight disinformation. {\color{pink} Add more examples to show that this problem is highlighted by different institutions with different/neutral political color.. }}

Section $230$ in the United States Communications Decency Act provides immunity for website platforms against the content created by users. Similar regulations exist in the European Union via the E-commerce Directive (2000) in articles $12$ and $15$.\footnote{See section $4$ in Bayer (2019)~\cite{Bayer} for a comprehensive overview of the mentioned articles.} Nevertheless, there is growing pressure for Mainstream Social Media Platforms (hereafter MSMP), such as Facebook, Twitter or Youtube, to moderate the available content. In particular, platforms take explicit actions when content is in violation of local laws in different jurisdictions, e.g. laws regarding defamation of a racial nature, dissemination of symbols from unconstitutional organizations, privacy protection, digital security, electoral laws. Facebook reports having implemented a total of $64.7$ thousand content restrictions based on local law across all countries in 2020.\footnote{See Facebook Transparency Center, Content restrictions based on Local Law: \href{https://transparency.fb.com/data/content\-restrictions}{transparency.fb.com/data/content\-restrictions}. We summed the count of content restrictions over all countries reported in the table, for $H1$ and $H2$ of the year 2020.} Google reports a total of $26$ thousand government requests to remove content from July 2020 to December 2020, among which $11.4$ thousand concerned Youtube.\footnote{See Google's Transparency report, government requests to remove content: \href{https://transparencyreport.google.com/government-removals/overview}{transparencyreport.google.com/government-removals/overview}.} Twitter reports having received $42.2$ thousand legal demands from third-parties from January to June 2020, and has responded by withholding $82$ thousand accounts and $3.1$ thousand tweets.\footnote{See Twitter Transparency website, Removal requests: \href{https://transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}{transparency.twitter.com/en/reports/removal-requests.html\#2020-jan-jun}.}

@@

Furthermore, MSMP are increasingly engaging in editorial tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and Youtube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include sub-categories such as violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam.\footnote{For an exhaustive overview of the community standards of Facebook, see: \href{https://www.facebook.com/communitystandards/}{facebook.com/communitystandards/}. For the Twitter  Rules see: \href{https://help.twitter.com/en/rules-and-policies/twitter-rules}{help.twitter.com/en/rules-and-policies/twitter-rules}, and for Youtube community guidelines see: \href{https://www.youtube.com/intl/en\_us/howyoutubeworks/policies/community\-guidelines/}{youtube.com/intl/en\_us/howyoutubeworks/policies/community\-guidelines/}.} While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries.

@@

During the COVID19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content.\footnote{For Facebook, Twitter and Youtube see respectively the following updates concerning COVID19 related misinformation policies: \href{https://support.google.com/youtube/answer/9891785}{https://support.google.com/youtube/answer/9891785}, \href{https://help.twitter.com/en/rules-and-policies/medical-misinformation-policy}{help.twitter.com/en/rules-and-policies/medical-misinformation-policy} and \href{https://support.google.com/youtube/answer/9891785}{support.google.com/youtube/answer/9891785}.} Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged\footnote{{\color{pink} put here the list of Facebook and Twitter terms about covid.}}, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of MSMP to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. In particular, policies regarding misinformation are not part of the set of platform rules or community guidelines\footnote{As of July 2021.}. The 2019 report of the Facebook Data Transparency Advisory Group\footnote{To read the full report, see: \href{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report?fbclid=IwAR2xMZr5GdD1GaNpjsXR3\_yeeIR4H9iFASfrni5HKcJVAO5oWA52bvwcZxU}{https://law.yale.edu/yls-today/news/facebook-data-transparency-advisory-group-releases-final-report?fbclid=IwAR2xMZr5GdD1GaNpjsXR3\_yeeIR4H9iFASfrni5HKcJVAO5oWA52bvwcZxU}.} (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report.}''. Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom.\footnote{See: \href{https://about.fb.com/news/2018/05/hard-questions-false-news/}{https://about.fb.com/news/2018/05/hard-questions-false-news/}.}. Similarly, Twitter communicates about actions related to misinformation via their Twitter Safety Blog/account\footnote{See \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{https://blog.twitter.com/en\_us/authors.TwitterSafety}.}. In a post on Youtube Official Blog\footnote{See \href{https://blog.youtube/inside-youtube/the-four-rs-of-responsibility-raise-and-reduce/}{https://blog.youtube/inside-youtube/the-four-rs-of-responsibility-raise-and-reduce/}.}, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation.

@@

In this article, we focus on MSMP's policies and actions regarding content with low credibility or false information, commonly referred to as {\it Fake News}.\footnote{For an overview on the {\it concept} of {\it Fake News}, we refer the reader to the article The science of fake news by Lazer et al. (2018)~\cite{lazer}.} The {\it Fake News} phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience. In particular, the intricacies of constructing a legal framework concerning misinformation and disinformation arises from the difficulty of identifying and qualifying a piece of online content as false or misleading, among an overwhelming quantity of daily produced content, without infringing existing laws.\footnote{See \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{A guide to anti-misinformation actions around the world} on the website of Poynter Institute.} The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society, in the rare event that it goes viral. This gave rise to a set of heterogenous fact-checking policies across mainstream platforms. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network.\footnote{For an overview of Facebook's third-party fact checking program see: \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}.} Facebook use a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section} Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for Youtube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels. \footnote{See \href{https://support.google.com/youtube/answer/9229632}{https://support.google.com/youtube/answer/9229632}.}

%{\color{pink} Note, put here something along the following lines: cite misinformation review article about support for fact-checking + hard to correct beliefs once it goes viral. cite Pennycook / Rand? }
%In particular, determining the truthfulness of content over controversial topics or ongoing political events proves to be a tricky task and correcting beliefs about viral content, fact-checked as false ... 
%For an overview of Youtube's policy on fact-checking see: \href{https://support.google.com/youtube/answer/9229632}{support.google.com/youtube/answer/9229632}. To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} 
%In particular, the peculiar task of determining the truthfulness of content over controversial topics or ongoing political events 
%\smallskip

%{\color{pink} Note, put here something along the following line: make point here about the fact that in the transparency centers we find things for the violations in the platforms rules but not misinformation.... needed so that the academic community can study the reach and impact of this phenomena and provide guidelines for adequate actions... + where to put this \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} ? }
%{\color{pink}  idea about academic community needs to be able to study this phenomenon and assess the impact of the policies, what effects they have etc. Violations fall into other categories: deceptive practices, impersonation, hate speech.  }

@@

Nevertheless, misinformation specific interventions by MSMP are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining MSMP's current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. For the purpose of clarity, we only focus on three platforms: Facebook, Twitter and Youtube. Both Facebook and Youtube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. More specifically, we survey a number of common policies used in order to tackle misinformation, across the three above cited MSMP, Facebook, Twitter and Youtube: temporary or permanent suspension of users, reducing the visibility of some content, introducing flags and notices. We do not provide  an exhaustive list of methods on how to investigate the platforms’ policies.\footnote{For a detailed list of interventions check the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}. } We rather provide a methodology to investigate key policies, that can be useful to researchers or journalists interested in implementing external monitoring. We summarize in table \ref{tab1} the tools used to collect the data from Facebook, Twitter and Youtube, that we use in multiple examples throughout the present article. We compile in Appendix \ref{links} in table \ref{summary}, URL links that redirect to the policies, regulations and transparency centers of Facebook, Twitter and Youtube. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.  %the announced policy for a specific user was implemented or not.

@@

When an account is permanently suspended by Facebook, it disappears from the platform. That is,  the data can no longer be scrapped and it also disappears from the CrowdTangle API.\footnote{CrowdTangle is a public insights tool owned and operated by Facebook, that exclusively tracks public content from Facebook public groups and pages.} Facebook publishes on monthly basis a {\it coordinated inauthentic behavior} report, where it informs how many personal accounts, pages or groups were deleted and to which {\it deceptive network} they may have belonged to.\footnote{See the \href{https://about.fb.com/news/2021/05/april-2021-coordinated-inauthentic-behavior-report/}{April 2021 report} as an example.} But as long as external persons do not have access to deleted accounts data, these reports cannot be verified by independent researchers or journalists.

Facebook can also apply a temporary suspension, and in this case the data can often be collected and analyzed. For example, \href{https://www.facebook.com/DonaldTrump/}{Donald Trump’s official Facebook page}  has been suspended following the Capitol attack on January 6, 2021.\footnote{See \href{https://www.facebook.com/zuck/posts/10112681480907401}{https://www.facebook.com/zuck/posts/10112681480907401}} Nevertheless the page’s data is still present in the CrowdTangle API. Thus, after manually adding this page to the CrowdTangle dashboard, we collected the $6$ $083$ posts it had published between January $1$, $2020$ and June $15$, $2021$ using the {\it posts} endpoint.\footnote{See the endpoint documentation for more details: \href{https://github.com/CrowdTangle/API/wiki/Posts}{https://github.com/CrowdTangle/API/wiki/Posts}.} We used Minet command line tool \cite{minet} to collect the data.\footnote{The exact command can be found \href{https://github.com/medialab/truth-and-trust-online-2021/blob/master/code/collect_facebook_crowdtangle_trump_data.sh}{here}.} We can verify on figure \ref{fig1_fb} that the {\it Donald J. Trump} page has not published any content since January $6$, 2021, and that this behavior is not consistent with the page’s previous activity: an average of $16$ posts were published each day on Facebook before the suspension.

@@

The website $globalresearch.ca$ is linked to the Twitter account {$@CRG\_CRM$}; which was recently suspended.\footnote{We noticed the message about the account suspension on May 25, 2021. But to the best of our knowledge, no official communication by Twitter has announced the suspension nor the exact date at which it was implemented. Hence the account may have gotten suspended anytime between April 15, 2021 and May 25, 2021 (see the suspension screenshot in panel $(a)$ of figure \ref{fig3}). Furthermore, $globalresearch.ca$ has multiple failed fact-checks, see \href{https://mediabiasfactcheck.com/global-research/}{https://mediabiasfactcheck.com/global-research/} and \href{https://feedback.news/media/AEKA}{https://feedback.news/media/AEKA}.} When a user searches via the twitter search-box for any URL link of this website, no results appear as shown in the screenshot in panel $(b)$ of figure \ref{fig3}, taken on June $14$, $2021$. To further investigate the possible implementation of a reduced visibility measure, we search via the Twitter API for tweets, excluding retweets, containing the query {\it globalresearch.ca} from January $1$, $2021$ until June $10$, $2021$.  As shown in panel (a) in figure \ref{fig4}, we find a strictly positive number of tweets containing the URL link {\it globalresearch.ca} throughout May $2021$ and the first week of June $2021$. Hence, the visibility of tweets containing this URL link has been reduced because users can no longer access tweets containing the URL link {\it globalresearch.ca} via the search box. Nevertheless users are not restrained from posting tweets containing this URL, as shown in the screenshot in panel $(a)$ of figure \ref{fig4bis}, found by taking the tweet ID of one of the collected tweets via the Twitter API. Furthermore, those collected Tweets have strictly positive engagement metrics as shown in panel $(b)$ of figure \ref{fig4}. Hence, the users who tweet articles from the {\it globalresearch.ca} website receive tweet level engagement from their own followers. Finally, when a user attempts to click on the URL link {\it globalresearch.ca} contained in the Tweet, a warning message appears and indicates that the link may be unsafe (see screenshot in panel (b) in figure \ref{fig4bis}).

@@

We further examine the placement of interstitials that indicate a possibly sensitive content. We find that only $2.97\%$ out of $323$ $938$ tweets containing a URL marked as False, have an interstitial ``potentially sensitive content''. In particular, many speak about COVID19 and do not contain a label to provide users with more context from authoritative sources. Figure \ref{fig_notice} provides an example of a tweet sharing a URL marked as False by a Fact-checker and which contains an interstitial ``potentially sensitive content''. We find 32 other Tweets, among our set of collected tweets, who share the exact same URL link as in the previous example. Among those 32 Tweets, only 5 tweets had the interstitial ``potentially sensitive content''. Again this points towards the non-automation of the placement of interstitials.\footnote{Notice that the appearance of interstitials may depend on the settings of a user's account and country or language specific regulations. In particular, in the settings of Twitter account, one can deactivate the display of interstitials for {\it sensitive content}, by ticking the box {\it Display media that may contain sensitive content} in the section {\it content you see}. Hence the numbers given are an upper bound of the number of interstitials indicating  possibly sensitive content.}

@@

Finally, Twitter Safety account announced in a Tweet\footnote{See Tweet ID : \href{https://twitter.com/TwitterSafety/status/1379515615954620418}{1379515615954620418}} on April 6, 2021 that their team ``will begin deploying automated tools to build on (their) efforts to label tweets that may contain misleading information around COVID-19 vaccinations". 
%we don't know what it means the articl xxx which was retracted by the journal and fack-check still appears in xx tweets with no label

@@

\subsection{Quick access to community guidelines and policies} \label{links}

\begin{landscape}

@@

\end{landscape}
%https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works
\end{document}
